---
title: "Predictive Modeling-Course Project"
author: "innuganti"
---
Data Citation:
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har 

# Overview
In this project, our goal is to predict the manner in which the samplers did the exercise. Used data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
```{r setup, include=FALSE}
library(ggplot2)
library(caret)
library(fscaret)
library(randomForest)
library(e1071)
library(lubridate)
library(scales)
library(RColorBrewer)
library(rattle)
library(rpart)
library(rpart.plot)
library(repmis)
```
Download Training and Testing Datasets
```{r}
set.seed(555)
url_training <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(url_training, destfile = "pml-training.csv")

url_testing <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(url_testing, destfile = "pml-testing.csv")
```
Read Data sets
```{r}
Training <- read.csv(url_training, na.strings=c("NA",""), header=TRUE)
dim(Training)

Testing <- read.csv(url_testing, na.strings=c("NA",""), header=TRUE)
dim(Testing)

```

Data Cleaning
Delete Predictors from the Training Dataset which contain missing values
```{r}
Training <- Training[, colSums(is.na(Training)) == 0]
Testing <- Testing[, colSums(is.na(Testing)) == 0]

# Remove columns which are unfit for prediction

Train <- Training[, -(1:7)]
Test <- Testing[, -(1:7)]

# Remaining Column Names
colnames(Train)
colnames(Test)
```

 Partitioning the Training Dataset 
```{r}
inTraining <- createDataPartition(Train$classe, p=0.7, list=FALSE)
Training_1 <- Train[inTraining, ]
Training_2 <- Train[-inTraining, ]
dim(Training_1); dim(Training_2)
```

Prediction with Decision Tree Modeling
```{r}

DT_Model <- rpart(classe ~ ., data=Training_1, method="class")
print(DT_Model, digits=3)
fancyRpartPlot(DT_Model)

```

```{r}
A_prediction <- predict(DT_Model, Training_1, type = "class")
cmtree <- confusionMatrix(A_prediction, Training_1$classe)
cmtree
```

Random Forest Modeling
```{r}
set.seed(555)
RF_Model <- randomForest(classe ~ ., data=Training_1)
B_Prediction <- predict(RF_Model, Training_1, type = "class")
cmrf <- confusionMatrix(B_Prediction, Training_1$classe)
cmrf

```

Results:
We were able to obtain 100% of accuracy with Random Forest model compared with Decision Tree model accuracy 74%.
Random Forest model will be chosen to predict the 20 test cases below.

Predicting Test Data
```{r}
B_Prediction <- predict(RF_Model, Test, type = "class")
B_Prediction
```
